---
title: "AGP OTU R analysis"
output: html_notebook
---


```{r}

devtools::load_all()
library(tidyverse)
library(dplyr)
library(qiime2R)
library(biomformat)
library(ggplot2)

```



## load metadata
```{r}

metadata = readr::read_tsv(system.file("data-raw/qiime/generated-files-20190512/ag.txt", package = "agp"))

country_info = readr::read_tsv(system.file(package="agp", "data-raw/countryInfo.txt"), comment = "#", col_names = FALSE, na = "") %>%
  select(X5,X9) %>%
  dplyr::rename(country="X5", continent="X9" )

```





## import biom and tax file

I have difficulities to import directly qza file either with `qiime2R::read_qza` either the biom file directly with `biomformat`.
For the moment, I extract qza with `unzip` within the shell and convert into tsv file with:

`biom convert -i feature-table.biom -o table.from_biom_w_tax.tsv --to-tsv --header-key taxonomy`



```{r eval=FALSE, include=FALSE}

biom_path = system.file("data-raw/qiime/generated-files-20190512/60049f6c-23e9-4bcd-8691-c9953d5b4a24/data/feature-table.biom", package="agp")
tax_path  = system.file("data-raw","qiime","generated-files-20190512","ag.fna.nobloom.min2.min1k.taxonomy.qza", package="agp")
#biom = qiime2R::read_qza(biom_path) # too slow!!
tax  = qiime2R::read_qza(tax_path)

biom = biomformat::read_biom("/home/tapju/storage/AmericanGut/agp/data-raw/qiime/generated-files-20190512/60049f6c-23e9-4bcd-8691-c9953d5b4a24/data/feature-table.biom")



```


### import OTU table
```{r}

#Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 4)

otu_path = system.file("data-raw/qiime/generated-files-20190512/60049f6c-23e9-4bcd-8691-c9953d5b4a24/data","table.from_biom_w_tax.tsv", package="agp")

#otu_path = system.file("data-raw/qiime/generated-files-small-set/a7fb8396-50b6-415b-8fee-4910e0d6d8b9/data","table.from_biom_w_tax.tsv", package="agp")

#tax_path  = system.file("data-raw","qiime","generated-files-20190512","ag.fna.nobloom.min2.min1k.taxonomy.qza", package="agp")


#otu = vroom::vroom(otu_path,comment = "", delim="\t", skip = 1)

otu = readr::read_tsv(otu_path,comment = "", skip=1) # very long!!
save(otu, file="otu.rds", compress=FALSE)
load(file="otu.rda")

```


```{r eval=FALSE, include=FALSE}
#otu_path = system.file("data-raw/qiime/generated-files-20190512/60049f6c-23e9-4bcd-8691-c9953d5b4a24/data","test.tsv", package="agp")
#otu_test = vroom::vroom(otu_path,comment = "", delim="\t", skip = 1)

```


### import tax table


```{r}

tax_path  = system.file("data-raw","qiime","generated-files-20190512","ag.fna.nobloom.min2.min1k.taxonomy.qza", package="agp")
tax  = qiime2R::read_qza(tax_path)

fna_path = system.file("data-raw","qiime","generated-files-20190512","ag.fna.nobloom.min2.min1k.qza", package="agp") 
fna      = qiime2R::read_qza(fna_path)



```

### pivot longer OTU table


```{r}

otu_longer_1 = 
otu[1:5000] %>% 
  #select(-taxonomy) %>%
  mutate_at(-1, ~na_if(.,0)) %>% 
  tidyr::pivot_longer(-`#OTU ID`, names_to = "sample_id", values_to = "count",    values_drop_na = TRUE) 

saveRDS(otu_longer_1, file = "otu_longer_1.rds")

otu_longer_2 = 
otu[c(1,5001:10000)] %>% 
  #select(-taxonomy) %>%
  mutate_at(-1, ~na_if(.,0)) %>% 
  tidyr::pivot_longer(-`#OTU ID`, names_to = "sample_id", values_to = "count",    values_drop_na = TRUE) 

saveRDS(otu_longer_2, file = "otu_longer_2.rds")

otu_longer_3 = 
otu[c(1,10001:15000)] %>% 
  #select(-taxonomy) %>%
  mutate_at(-1, ~na_if(.,0)) %>% 
  tidyr::pivot_longer(-`#OTU ID`, names_to = "sample_id", values_to = "count",    values_drop_na = TRUE) 

saveRDS(otu_longer_3, file = "otu_longer_3.rds")

last_sample=dim(otu)[2]-1

otu_longer_4 = 
otu[c(1,15001:last_sample)] %>% 
  #select(-taxonomy) %>%
  mutate_at(-1, ~na_if(.,0)) %>% 
  tidyr::pivot_longer(-`#OTU ID`, names_to = "sample_id", values_to = "count",    values_drop_na = TRUE) 

saveRDS(otu_longer_4, file = "otu_longer_4.rds")


otu_longer_all = rbind(otu_longer,otu_longer_2,otu_longer_3,otu_longer_4)

saveRDS(otu_longer_all, file = "otu_longer_all.rds")



```



### focus on bifidobacterium


```{r}

tax$data %>% 
  #mutate(Taxon = Taxon %>% as.character) %>%
  filter(Taxon %in% grep("Bifidobacterium",Taxon,value = TRUE), Confidence > 0.90) %>% pull(Feature.ID) -> bif_otu



otu_longer_all %>%
  #filter(sample_id %in% c("10317.000042798","10317.000058805")) %>%
  group_by(sample_id) %>%
  mutate(count = SpiecEasi::clr(count)) -> otu_clr

otu_clr %>%
  filter(`#OTU ID` %in% bif_otu) %>%
  ungroup() %>%
  group_by(`#OTU ID`) %>%
  summarise(q25=quantile(count,0.25),q50=quantile(count,0.50),q75=quantile(count,0.75), n=n() ) -> bif_stat_summary

  
  
```



```{r}

bif_stat_summary %>%
  arrange(desc(n)) %>%
  ggplot() + geom_point(aes(x=n,y=q50)) + 
  geom_point(aes(x=n,y=q25), col="red") + 
  geom_point(aes(x=n,y=q75), col="red") +
  scale_x_log10()

bif_stat_summary %>%
  filter(n>10) %>%
  merge(tax$data, by.x = "#OTU ID", by.y="Feature.ID") %>%
  arrange(desc(n)) %>% pull(`#OTU ID`) -> prevalent_bif



```



```{r}

metadata %>%
  select("#SampleID","age_cat","age_corrected","sex","country_of_birth","country") %>%
  merge(.,otu_clr, by.x="#SampleID", by.y="sample_id") -> otu_metadata

otu_metadata %>%
  filter(`#OTU ID` %in% bif_otu) %>%
  group_by(`#OTU ID`,age_cat) %>%
  summarise(q25=quantile(count,0.25),q50=quantile(count,0.50),q75=quantile(count,0.75), n=n() ) -> bif_stat_summary_by_age


metadata %>%
  select("#SampleID","age_cat","age_corrected","sex","country_of_birth","country") %>%
  group_by(age_cat) %>%
  filter(age_cat != "baby") %>%
  filter(!is.na(age_cat), age_cat!="Not provided") %>%
  sample_n(size=300) %>% pull(`#SampleID`) -> sample_sample_id


otu_metadata %>%
  filter(`#SampleID` %in% sample_sample_id) %>%
  filter(`#OTU ID` %in% bif_otu) %>%
  group_by(`#OTU ID`,age_cat) %>%
  summarise(q25=quantile(count,0.25),q50=quantile(count,0.50),q75=quantile(count,0.75), n=n() ) %>%
  ungroup() %>%
  select(`#OTU ID`, age_cat) %>%
  group_by(age_cat) %>%
  summarise(n=n())



otu_metadata %>% 
  filter(`#OTU ID` %in% bif_otu) %>%
  group_by(`#SampleID`,age_cat) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  group_by(age_cat) %>%
  summarise(q25=quantile(n,0.25),q50=quantile(n,0.50),q75=quantile(n,0.75))



```



```{r fig.height=10, fig.width=10}
otu_clr %>% pull(sample_id) %>% unique %>% length -> nb_samples

bif_stat_summary_by_age %>% 
  mutate(p=n/nb_samples) %>%
  filter(!is.na(age_cat), age_cat!="Not provided") %>%
  ungroup() %>%
  mutate(age_cat=as.factor(age_cat)) %>% 
  filter(`#OTU ID` %in% prevalent_bif) %>%
  mutate(age_cat=age_cat %>% forcats::fct_relevel("baby","child","teen",after=0)) %>%
  ggplot() + geom_point(aes(x=age_cat,y=p)) + facet_wrap(~`#OTU ID`, scale="free_y", ncol=3)
  

bif_stat_summary_by_age %>% 
  mutate(p=n/nb_samples) %>%
  filter(!is.na(age_cat), age_cat!="Not provided") %>%
  ungroup() %>%
  mutate(age_cat=as.factor(age_cat)) %>% 
  filter(`#OTU ID` %in% prevalent_bif) %>%
  mutate(age_cat=age_cat %>% forcats::fct_relevel("baby","child","teen",after=0)) %>%
  ggplot() + geom_line(aes(x=age_cat,y=q50, group=`#OTU ID`)) + facet_wrap(~`#OTU ID`, scale="free_y", ncol=3)


bif_stat_summary_by_age %>% 
  mutate(p=n/nb_samples) %>%
  filter(!is.na(age_cat), age_cat!="Not provided") %>%
  ungroup() %>%
  mutate(age_cat=as.factor(age_cat)) %>% 
  filter(`#OTU ID` %in% prevalent_bif) %>%
  mutate(age_cat=age_cat %>% forcats::fct_relevel("baby","child","teen",after=0)) %>%
  ggplot() + geom_line(aes(x=age_cat,y=q50, group=`#OTU ID`)) + facet_wrap(~`#OTU ID`, scale="free_y", ncol=3)




```


```{r}



fna$data[names(fna$data) %in% bif_otu] %>%
  ape::as.DNAbin() -> fna_dnabin
  
  ape::dist.dna(fna_dnabin, model = "raw") -> fna_dist
  

  
  


bif_clust = fpc::kmeansCBI(krange=4, data = fna_dist, criterion="ch")



bif_phylo_pcoa = ade4::dudi.pco(sqrt(fna_dist), scannf=FALSE, nf=3)

bif_phylo_pcoa$li %>%
  merge(bif_stat_summary, by.x="row.names", by.y="#OTU ID") %>%
  merge(as.matrix(bif_clust$partition) %>% as.data.frame(), by.x="Row.names", by.y="row.names") %>%
  mutate(cut_n=cut(log10(n),4)) %>%
  ggplot() + geom_point(aes(x=A1,y=A2, col=as.character(V1))) + facet_wrap(~cut_n)

cumsum(bif_phylo_pcoa$eig/sum(bif_phylo_pcoa$eig))


```

### classify reads using decipher
```{r}
library("DECIPHER")

load("~/SILVA_SSU_r132_March2018.RData")
load("~/Contax_v1_March2018.RData")
load("~/GTDB_r89-mod_June2019.RData")

```


```{r}

fna$data[names(fna$data) %in% bif_otu] -> bif_DNAStringSet
dna <- bif_DNAStringSet
ids <- IdTaxa(dna, trainingSet, strand="top", processors=NULL, verbose=FALSE) # use all processors
ranks <- c("domain", "phylum", "class", "order", "family", "genus", "species") # ranks of interest
# Convert the output object of class "Taxa" to a matrix analogous to the output from assignTaxonomy
taxid <- t(sapply(ids, function(x) {
        m <- match(ranks, x$rank)
        taxa <- x$taxon[m]
        #taxa[startsWith(taxa, "unclassified_")] <- NA
        taxa
}))

colnames(taxid) = ranks

taxid %>%
  as.data.frame() %>%
  filter(is.na(genus)) 

taxid %>%
  as.data.frame() %>%
  unique()

```



```{r}

bif_phylo_pcoa$li %>%
  merge(bif_stat_summary, by.x="row.names", by.y="#OTU ID") %>%
  merge(as.matrix(bif_clust$partition) %>% as.data.frame(), by.x="Row.names", by.y="row.names") %>%
  merge(taxid %>% as.data.frame, by.x="Row.names", by.y="row.names") %>%
  mutate(cut_n=cut(log10(n),4)) %>%
  ggplot() + geom_point(aes(x=A1,y=A2, col=as.character(V1))) + facet_grid(genus~cut_n)


```

