---
title: "curated metagenomics data V3 enterotying"
output: html_notebook
---

```{r}
library(purrr)
library(dplyr)


```



```{r}

## group all bootstrap within one file
fit_genus_bootstrap_curated = vector("list",2)
load("fit_genus_list_1_curated_v3.rda")
fit_genus_bootstrap_curated[[1]] = fit_genus_list
load("fit_genus_list_2_curated_v3.rda")
fit_genus_bootstrap_curated[[2]] = fit_genus_list
load("fit_genus_list_3_curated_v3.rda")
fit_genus_bootstrap_curated[[3]] = fit_genus_list
load("fit_genus_list_4_curated_v3.rda")
fit_genus_bootstrap_curated[[4]] = fit_genus_list
load("fit_genus_list_5_curated_v3.rda")
fit_genus_bootstrap_curated[[5]] = fit_genus_list

```

extract BIC and Laplace coefficients from boostrap models  

```{r}
BIC_list = map(fit_genus_bootstrap_curated, function(z) map(z, function(y) sapply(y, function(x){attr(x,"goodnessOfFit")[["BIC"]]}))) 
lplc_list = map(fit_genus_bootstrap_curated, function(z) map(z, function(y) sapply(y, function(x){attr(x,"goodnessOfFit")[["Laplace"]]})))
k = map(fit_genus_bootstrap_curated, function(z) map(z, function(y) sapply(y, function(x){attr(x,"mixture")$Weight %>% length})))[[1]][[1]]


BIC_df = NULL

for(i in 1:5) {
  for(j in 1:5){
    
    BIC_df=cbind(BIC_df,BIC_list[[i]][[j]])
    
  }
}



lplc_df = NULL

for(i in 1:5) {
  for(j in 1:5){
    
    lplc_df=cbind(lplc_df,lplc_list[[i]][[j]])
    
  }
}
```


plot BIC and laplace results
```{r fig.height=6, fig.width=12}

nb_seed=5

cowplot::plot_grid(

BIC_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:nb_seed, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*nb_seed))) %>%
  ggplot() + geom_line(aes(x=`Dirichlet component`,y=value,group=variable,color=boot)) +
  ggtitle("BIC optimun") + #xlab("Dirichlet component") +
  theme_bw() + scale_x_continuous(limits = c(1,101), breaks = scales::pretty_breaks(n = 20)),

lplc_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:nb_seed, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*nb_seed))) %>%
  ggplot() + geom_line(aes(x=`Dirichlet component`,y=value,group=variable,color=boot)) +
  ggtitle("Laplace optimun") + #xlab("Dirichlet component") +
  theme_bw() + scale_x_continuous(limits = c(1,101), breaks = scales::pretty_breaks(n = 20)),



BIC_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:nb_seed, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*nb_seed))) %>%
  group_by(boot,seed) %>%
  filter(value==min(value)) %>%
  ggplot() + geom_bar(aes(x=`Dirichlet component`)) + 
  ggtitle("BIC optimun") + 
  xlab("Dirichlet components")+
  theme_bw() + scale_x_continuous(limits = c(1,30), breaks = scales::pretty_breaks(n = 30)),


lplc_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:nb_seed, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*nb_seed))) %>%
  group_by(boot,seed) %>%
  filter(value==min(value)) %>%
  ggplot() + geom_bar(aes(x=`Dirichlet component`)) + 
  ggtitle("Laplace optimun") + 
  xlim(1,30) +  xlab("Dirichlet components")+
  theme_bw() + scale_x_continuous(limits = c(1,30), breaks = scales::pretty_breaks(n = 30)),
labels = "AUTO"
)

ggsave(filename = "figures/figS3.pdf")

```

BIC and Laplace have a global optimun among differents seeds and bootstraps. All bootstraps found at least one model optimal fit at k=24. Majority of best model based on Laplace were k=24. For BIC, k=24 was also found as an optimum. For downstream, analysis we then choose k=24.

```{r}

boot=1
seed=1
fit_genus_list = fit_genus_bootstrap_curated[[boot]]
best_genus_lplc = 24 
best_genus_lplc_idx = which(sapply(fit_genus_list[[seed]], function(x) attr(x,"mixture")$Weight %>% length)==best_genus_lplc) # 19th items from list corresponded to k=24 here


enterotypes_curated =

  fit_genus_list[[seed]][[best_genus_lplc_idx]] %>%
  mixture(assign=TRUE) %>% as.data.frame %>% set_colnames(c("Enterotypes_id"))







```



```{r fig.height=6, fig.width=14}


###################################################
### code chunk number 11: posterior-mean-diff
###################################################
p0 <- fitted(fit_genus_list[[seed]][[1]], scale=TRUE)     # scale by theta
pbest <- fitted(fit_genus_list[[seed]][[best_genus_lplc_idx]], scale=TRUE)
colnames(pbest) <- paste("m", 1:best_genus_lplc, sep="")
(meandiff <- colSums(abs(pbest - as.vector(p0))))
sum(meandiff)


###################################################
### code chunk number 12: table-1
###################################################
diff <- rowSums(abs(pbest - as.vector(p0)))
o <- order(diff, decreasing=TRUE)
cdiff <- cumsum(diff[o]) / sum(diff)
df <- head(cbind(Mean=p0[o], pbest[o,], diff=diff[o], cdiff), 30)
df %>% as.data.frame()

df_curated_v3 = df %>% as.data.frame()


m_order = df %>% as.data.frame() %>% select(-Mean, -diff,-cdiff)  %>% BiotypeR::dist.JSD() %>% hclust(method="complete") %>% unclass %>% .$order
#m_order = df %>% as.data.frame() %>% select(-Mean, -diff,-cdiff)  %>% t %>% dist() %>% hclust(method="complete") %>% unclass %>% .$order 


df %>% as.data.frame() %>%
  tibble::rownames_to_column("taxa") %>%
  mutate(taxa=factor(taxa, levels=row.names(df))) %>% 
  reshape2::melt(id.vars="taxa") %>%
  filter(!variable %in% c("Mean","diff","cdiff")) %>%
  mutate(variable = variable %>% as.character) %>% 
  mutate(variable = variable %>% factor(levels=paste0("m",1:24)[m_order]) )  -> genus_alpha_weight_curated

save(genus_alpha_weight_curated, file="genus_alpha_weight_curated.rda")


df %>% as.data.frame() %>%
  tibble::rownames_to_column("taxa") %>%
  mutate(taxa=factor(taxa, levels=row.names(df))) %>% 
  reshape2::melt(id.vars="taxa") %>%
  filter(variable %in% c("Mean","diff","cdiff"))


df %>% as.data.frame() %>%
  tibble::rownames_to_column("taxa") %>%
  mutate(taxa=factor(taxa, levels=row.names(df))) %>% 
  reshape2::melt(id.vars="taxa") %>%
  filter(!variable %in% c("Mean","diff","cdiff")) %>%
  mutate(variable = variable %>% as.character) %>% 
  mutate(variable = variable %>% factor(levels=paste0("m",1:24)[m_order]) ) %>% #pull(variable)
  #group_by(taxa) %>%
  #mutate(total = sum(value)) %>%
  #mutate(value=value/total) %>%
  ggplot() + 
  geom_tile(aes(x=variable,y=taxa,fill=log10(value))) + 
  scale_fill_viridis_c("DMM model alpha") +
  ylab("taxa") +
  xlab("DMM components")
  
  
```


### correlation between DMM alpha_weight and shannon alpha diversity

```{r}






df %>% as.data.frame() %>%
  tibble::rownames_to_column("taxa") %>%
  mutate(taxa=factor(taxa, levels=row.names(df))) %>% 
  reshape2::melt(id.vars="taxa") %>%
  filter(!variable %in% c("Mean","diff","cdiff")) %>%
  mutate(variable = variable %>% as.character) %>% 
  mutate(variable = variable %>% factor(levels=paste0("m",1:24)[m_order]) ) %>%
  
  merge(
    
    shannon_curated_v3 %>%
  merge(enterotypes_curated, by="row.names") %>%
  mutate(Enterotypes_id=paste0("m",Enterotypes_id)) %>%
  group_by(Enterotypes_id) %>%
  summarise(m=median(shannon)),
  
  by.x="variable", by.y="Enterotypes_id"
    
    
    
  ) %>%
  group_by(taxa) %>%
  do(broom::tidy(cor.test(.$value, .$m, method="spearman"))) %>%
  ungroup() %>%
  mutate(p.value = p.adjust(p.value, method = "fdr") %>% round(4)) %>%
  arrange(desc(abs(estimate))) %>%
  select(taxa,estimate,p.value)-> cor_alpha_table
  
write.csv2(cor_alpha_table, file="cor_alpha_table.csv")



# df_curated_v3 %>%
#   select(Mean,diff) %>%
#   merge(cor_alpha_table,.,by.x="taxa", by.y="row.names") %>%
#   ggplot(aes(x=abs(estimate),y=diff)) + geom_point()
#   #ggrepel::geom_label_repel(aes(label=taxa), max.overlaps = 20)


```





look at theta per component
```{r}

#fitted(fit_genus_list[[1]][[best_genus_lplc]])

mixturewt(fit_genus_list[[seed]][[1]])

mixturewt(fit_genus_list[[seed]][[best_genus_lplc_idx]])

mixturewt(fit_genus_list[[seed]][[best_genus_lplc_idx]]) %>% 
  tibble::rowid_to_column("k") %>% 
  ggplot() + 
  geom_bar(aes(x=k,y=theta),stat="identity") + 
  geom_hline(yintercept = mixturewt(fit_genus_list[[seed]][[1]])$theta, col="red") + 
  xlab("DMM community types") + ylab("theta extracted from DMM modelling") +
  scale_x_continuous(breaks = 1:24) +
  cowplot::theme_cowplot() -> figS4A
  
figS4A  

```






```{r fig.height=8, fig.width=15}

df %>% as.data.frame() %>%
  select(-diff,-cdiff,-Mean) %>%
  mutate_at(1:19,~log10(.)) %>%
  t %>%
  as.data.frame %>%
  ade4::dudi.pca(scannf = F,nf=3) %>% .$li %>% ade4::s.label()




df %>% as.data.frame() %>%
  select(-diff,-cdiff,-Mean) %>%
  tibble::rownames_to_column("taxa") %>%
  #mutate(taxa2=taxa) %>%
  #mutate(taxa2=ifelse())
    as_tibble %>%
  .[c(1:3,8,11,30),] %>%
  reshape2::melt(id.vars="taxa") %>%
  group_by(variable) %>%
  mutate(ymax=cumsum(value)) %>%
  mutate(ymin=c(0, head(ymax, n=-1))) %>%
  #ggplot() + geom_bar(aes(y=value,x=variable,fill=taxa),stat="identity") 
  ggplot(aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=taxa)) + 
  geom_rect() +      
  coord_polar(theta="y") + 
  facet_wrap(~variable) + 
  xlim(c(2, 4)) + theme_void() +
  scale_fill_brewer("",type="qual")

ggsave("curated_v3_microbiota_types.pdf")


```




### classify data with the choosen model


```{r}
load("curated_v3_genus_count.rda")


predict(fit_genus_list[[1]][[best_genus_lplc_idx]], t(curated_v3_genus_count), assign = TRUE) %>%
  reshape2::melt() %>%
  group_by(Var1) %>%
  filter(value==max(value)) %>%
  dplyr::rename(`sample_name`="Var1", Enterotypes_id="Var2") %>%
  mutate(set=ifelse(sample_name %in% row.names(attr(fit_genus_list[[1]][[best_genus_lplc_idx]],"group")), "train", "predicted")) -> enterotypes_prediction

table(enterotypes_prediction$value > 0.50) %>% prop.table()
table(enterotypes_prediction$value > 0.80) %>% prop.table()

enterotypes_prediction %>%
  ggplot() + geom_histogram(aes(x=value), bins = 50) + 
  #facet_wrap(~set) + 
  scale_x_continuous(limits = c(0,1.05), labels = scales::percent) +
  xlab("classification confidence") +
  ylab("number of samples") +
  cowplot::theme_cowplot() -> figS4B

figS4B


```


```{r fig.height=4, fig.width=10}


cowplot::plot_grid(figS4A,figS4B, labels = "AUTO", rel_widths = c(1.7,1))

ggsave("figures/figS4.pdf", h=5, w=10)


```



```{r}

enterotypes_prediction %>%
  #as.data.frame %>%
  #rbind(data.frame(sample_name = outliers,Enterotypes_id=NA,value=NA,set="outliers", stringsAsFactors = FALSE)) %>%
  write.csv2(file="enterotypes_curated_v3_prediction.csv")






```








```{r fig.height=5, fig.width=10}

row.names(df_gutzilla) = sapply(row.names(df_gutzilla) %>% strsplit(split="g__"), function(x) ifelse(is.na(x[2]), paste0(sample(LETTERS, 1),sample(LETTERS, 1)), x[2]))

colnames(df_gutzilla) = colnames(df_gutzilla) %>% gsub("m","M", .) 

merge(df_curated_v3 %>% select(-m19,-m22,-m23,-m24),df_gutzilla, by="row.names") %>% 
  select(contains("m",ignore.case = TRUE)) %>% 
  select(-Mean.x,-Mean.y, -Row.names) %>% 
  colnames() %>%
  tibble::tibble(Microbiota=.) %>%
  mutate(dataset=ifelse(Microbiota %>% grepl("\\M", .), "AGP","CuratedMGdb")) %>%
  #mutate(random=rnorm(18+19)) %>%
  #merge(microbiota_types_alpha_diversity, by.x="Microbiota", by.y="Enterotypes_id") %>%
  #select(-rank, -Shannon) %>%
  #select(-rank) %>%
  #dplyr::rename(`Shannon scaled`="scale") %>%
  #merge(diet_coa, by.x="Microbiota", by.y ="row.names", all = TRUE ) %>%
  #select(-Axis2) %>%
  #dplyr::rename
  tibble::column_to_rownames("Microbiota") -> microbiota_annotation


merge(df_curated_v3%>% select(-m19,-m22,-m23,-m24),df_gutzilla, by="row.names") %>% 
  select(contains("m",ignore.case = TRUE)) %>% 
  select(-Mean.x,-Mean.y) %>%
  #mutate(Row.names = Row.names %>% str_split_fixed(pattern =";",  n = 6) %>% .[,6] %>% gsub("g__","",.)) %>%
  tibble::column_to_rownames("Row.names") %>% as.matrix() -> microbiota_types

microbiota_types_jsd = Rfast::Dist(t(microbiota_types)) 
colnames(microbiota_types_jsd) = row.names(microbiota_types_jsd) = colnames(microbiota_types)

microbiota_types_jsd = microbiota_types_jsd %>% as.dist() %>% sqrt()

ann_colors = list(
    dataset = c(AGP="pink", CuratedMGdb="lightblue"),
    `Shannon scaled` = RColorBrewer::brewer.pal(4, name="OrRd"),
    Axis1 = RColorBrewer::brewer.pal(4, name="BrBG"),
    Axis2 = RColorBrewer::brewer.pal(4, name="BrBG")
)


  #as.matrix %>% #log10() %>%
  pheatmap::pheatmap(mat=log10(microbiota_types), cutree_cols = 7, annotation_col = microbiota_annotation, cellwidth = 10, cellheight = 20, clustering_distance_cols = microbiota_types_jsd, annotation_colors = ann_colors, color=viridis::viridis_pal()(100)) %>% ggplotify::as.ggplot() -> fig3A
fig3A
ggsave("dmm_types_clustering_curated_agp.pdf")

```

### permanova en AGP/curated dmm partitions

```{r}
anova(vegan::betadisper(microbiota_types_jsd, microbiota_annotation$dataset ))


vegan::adonis(microbiota_types_jsd~microbiota_annotation$dataset)


```



