---
title: "curated metagenomics data V3 enterotying"
output: html_notebook
---

```{r}
library(purrr)
library(dplyr)


```



```{r}

## group all bootstrap within one file
fit_genus_bootstrap_curated = vector("list",2)
load("fit_genus_list_1_curated_v3.rda")
fit_genus_bootstrap_curated[[1]] = fit_genus_list
load("fit_genus_list_2_curated_v3.rda")
fit_genus_bootstrap_curated[[2]] = fit_genus_list
load("fit_genus_list_3_curated_v3.rda")
fit_genus_bootstrap_curated[[3]] = fit_genus_list
load("fit_genus_list_4_curated_v3.rda")
fit_genus_bootstrap_curated[[4]] = fit_genus_list
load("fit_genus_list_5_curated_v3.rda")
fit_genus_bootstrap_curated[[5]] = fit_genus_list

```

extract BIC and Laplace coefficients from boostrap models  

```{r}
BIC_list = map(fit_genus_bootstrap_curated, function(z) map(z, function(y) sapply(y, function(x){attr(x,"goodnessOfFit")[["BIC"]]}))) 
lplc_list = map(fit_genus_bootstrap_curated, function(z) map(z, function(y) sapply(y, function(x){attr(x,"goodnessOfFit")[["Laplace"]]})))
k = map(fit_genus_bootstrap, function(z) map(z, function(y) sapply(y, function(x){attr(x,"mixture")$Weight %>% length})))[[1]][[1]]


BIC_df = NULL

for(i in 1:5) {
  for(j in 1:5){
    
    BIC_df=cbind(BIC_df,BIC_list[[i]][[j]])
    
  }
}



lplc_df = NULL

for(i in 1:5) {
  for(j in 1:5){
    
    lplc_df=cbind(lplc_df,lplc_list[[i]][[j]])
    
  }
}
```


plot BIC and laplace results
```{r fig.height=6, fig.width=12}

nb_seed=5

cowplot::plot_grid(

BIC_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:nb_seed, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*nb_seed))) %>%
  ggplot() + geom_line(aes(x=`Dirichlet component`,y=value,group=variable,color=boot)) +
  ggtitle("BIC optimun") + #xlab("Dirichlet component") +
  theme_bw() + scale_x_continuous(limits = c(1,101), breaks = scales::pretty_breaks(n = 20)),

lplc_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:nb_seed, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*nb_seed))) %>%
  ggplot() + geom_line(aes(x=`Dirichlet component`,y=value,group=variable,color=boot)) +
  ggtitle("Laplace optimun") + #xlab("Dirichlet component") +
  theme_bw() + scale_x_continuous(limits = c(1,101), breaks = scales::pretty_breaks(n = 20)),



BIC_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:nb_seed, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*nb_seed))) %>%
  group_by(boot,seed) %>%
  filter(value==min(value)) %>%
  ggplot() + geom_bar(aes(x=`Dirichlet component`)) + 
  ggtitle("BIC optimun") + 
  xlab("Dirichlet components")+
  theme_bw() + scale_x_continuous(limits = c(1,30), breaks = scales::pretty_breaks(n = 30)),


lplc_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:nb_seed, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*nb_seed))) %>%
  group_by(boot,seed) %>%
  filter(value==min(value)) %>%
  ggplot() + geom_bar(aes(x=`Dirichlet component`)) + 
  ggtitle("Laplace optimun") + 
  xlim(1,30) +  xlab("Dirichlet components")+
  theme_bw() + scale_x_continuous(limits = c(1,30), breaks = scales::pretty_breaks(n = 30)),
labels = "AUTO"
)



```

BIC and Laplace have a global optimun among differents seeds and bootstraps. All bootstraps found at least one model optimal fit at k=24. Majority of best model based on Laplace were k=24. For BIC, k=24 was also found as an optimum. For downstream, analysis we then choose k=24.

```{r}

boot=1
seed=1
fit_genus_list = fit_genus_bootstrap_curated[[boot]]
best_genus_lplc = 24 
best_genus_lplc_idx = which(sapply(fit_genus_list[[seed]], function(x) attr(x,"mixture")$Weight %>% length)==best_genus_lplc) # 19th items from list corresponded to k=24 here


enterotypes_curated =

  fit_genus_list[[seed]][[best_genus_lplc_idx]] %>%
  mixture(assign=TRUE) %>% as.data.frame %>% set_colnames(c("Enterotypes_id"))







```



```{r fig.height=6, fig.width=14}
###################################################
### code chunk number 11: posterior-mean-diff
###################################################
p0 <- fitted(fit_genus_list[[seed]][[1]], scale=TRUE)     # scale by theta
pbest <- fitted(fit_genus_list[[seed]][[best_genus_lplc_idx]], scale=TRUE)
colnames(pbest) <- paste("m", 1:best_genus_lplc, sep="")
(meandiff <- colSums(abs(pbest - as.vector(p0))))
sum(meandiff)


###################################################
### code chunk number 12: table-1
###################################################
diff <- rowSums(abs(pbest - as.vector(p0)))
o <- order(diff, decreasing=TRUE)
cdiff <- cumsum(diff[o]) / sum(diff)
df <- head(cbind(Mean=p0[o], pbest[o,], diff=diff[o], cdiff), 30)
df %>% as.data.frame()

df_gutzilla = df %>% as.data.frame()


m_order = df %>% as.data.frame() %>% select(-Mean, -diff,-cdiff) %>% t %>% dist() %>% hclust(method="complete") %>% unclass %>% .$order
 

df %>% as.data.frame() %>%
  tibble::rownames_to_column("taxa") %>%
  mutate(taxa=factor(taxa, levels=row.names(df))) %>% 
  reshape2::melt(id.vars="taxa") %>%
  filter(!variable %in% c("Mean","diff","cdiff")) %>%
  mutate(variable = variable %>% factor(levels=paste0("m",1:24)[m_order]) ) %>%
  #group_by(taxa) %>%
  #mutate(total = sum(value)) %>%
  #mutate(value=value/total) %>%
  ggplot() + 
  geom_tile(aes(x=variable,y=taxa,fill=log10(value))) + 
  scale_fill_viridis_c("DMM model alpha") +
  ylab("taxa") +
  xlab("DMM components")
  
```







look at theta per component
```{r}

#fitted(fit_genus_list[[1]][[best_genus_lplc]])

mixturewt(fit_genus_list[[seed]][[1]])

mixturewt(fit_genus_list[[seed]][[best_genus_lplc_idx]])

mixturewt(fit_genus_list[[seed]][[best_genus_lplc_idx]]) %>% 
  tibble::rowid_to_column("k") %>% 
  ggplot() + 
  geom_bar(aes(x=k,y=theta),stat="identity") + 
  geom_hline(yintercept = mixturewt(fit_genus_list[[seed]][[1]])$theta, col="red")

```






```{r fig.height=8, fig.width=15}

df %>% as.data.frame() %>%
  select(-diff,-cdiff,-Mean) %>%
  mutate_at(1:19,~log10(.)) %>%
  t %>%
  as.data.frame %>%
  ade4::dudi.pca(scannf = F,nf=3) %>% .$li %>% ade4::s.label()




df %>% as.data.frame() %>%
  select(-diff,-cdiff,-Mean) %>%
  tibble::rownames_to_column("taxa") %>%
  #mutate(taxa2=taxa) %>%
  #mutate(taxa2=ifelse())
    as_tibble %>%
  .[c(1:3,5,29,4),] %>%
  reshape2::melt(id.vars="taxa") %>%
  group_by(variable) %>%
  mutate(ymax=cumsum(value)) %>%
  mutate(ymin=c(0, head(ymax, n=-1))) %>%
  #ggplot() + geom_bar(aes(y=value,x=variable,fill=taxa),stat="identity") 
  ggplot(aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=taxa)) + 
  geom_rect() +      
  coord_polar(theta="y") + 
  facet_wrap(~variable) + 
  xlim(c(2, 4)) + theme_void() +
  scale_fill_brewer("",type="qual")

ggsave("curated_v3_microbiota_types.pdf")


```




### classify data with the choosen model


```{r}
load("curated_v3_genus_count.rda")


predict(fit_genus_list[[1]][[best_genus_lplc_idx]], t(curated_v3_genus_count), assign = TRUE) %>%
  reshape2::melt() %>%
  group_by(Var1) %>%
  filter(value==max(value)) %>%
  dplyr::rename(`sample_name`="Var1", Enterotypes_id="Var2") %>%
  mutate(set=ifelse(sample_name %in% row.names(attr(fit_genus_list[[1]][[best_genus_lplc_idx]],"group")), "train", "predicted")) -> enterotypes_prediction


enterotypes_prediction %>%
  ggplot() + geom_histogram(aes(x=value, fill=value>0.80), bins = 50) + 
  #facet_wrap(~set) + 
  scale_x_continuous(limits = c(0,1.05), labels = scales::percent)


```





```{r}

enterotypes_prediction %>%
  #as.data.frame %>%
  #rbind(data.frame(sample_name = outliers,Enterotypes_id=NA,value=NA,set="outliers", stringsAsFactors = FALSE)) %>%
  write.csv2(file="enterotypes_curated_v3_prediction.csv")






```













