---
title: "hierarchical enterotyping"
output: html_notebook
---



load genus abundance
```{r}

genus_path = system.file("data-raw/qiime/generated-files-20190512/taxa/genus.qza", package = "gutzilla")

genus = qiime2R::read_qza(genus_path)$data %>% as.data.frame %>% tibble::rownames_to_column("taxa")  %>% as_tibble()


```

load metadata (cleaned) and add united nation subregions
```{r}

data("UNSD_countries")

metadata = readr::read_csv2(system.file("data-raw/Metadata_10317_20191022-112414_curatedv4_VSv1.csv", package = "agp"))

#to do: check issue parsing IBD variables

colnames(metadata) = stringr::str_to_lower(colnames(metadata))



metadata %>%
  select(country_of_birth) %>%
  merge(UNSD_countries %>% 
          select(`Country or Area`, `Sub-region Name`, `Region Name`) %>% 
          mutate(`Country or Area` =  gsub("United Kingdom of Great Britain and Northern Ireland", "United Kingdom", `Country or Area`)) %>% 
          mutate(`Country or Area` =  gsub("United States of America", "USA", `Country or Area`)) %>%
          mutate(`Sub-region Name` =  ifelse(`Sub-region Name` %in% c("Central Asia", "Southern Asia"), "Central and Southern Asia",`Sub-region Name`)) %>%
          mutate(`Sub-region Name` =  ifelse(`Sub-region Name` %in% c("Melanesia", "Micronesia","Polynesia"), "Oceania (others)",`Sub-region Name`)), 
        by.x="country_of_birth", by.y="Country or Area") %>% 
  group_by(`Region Name`,`Sub-region Name`) %>% summarise(n=n()) %>%
  ungroup() %>%
  ggplot() + geom_bar(aes(y=n, x=`Sub-region Name`), stat="identity") + scale_y_log10() +
  facet_wrap(~`Region Name`, scales = "free_x", nr=1) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  xlab("Region of Birth\n(United Nations-m49)") +
  ylab("Number of stool sample")
  
ggsave("Region_of_birth.pdf")

metadata <- metadata %>%
  merge(UNSD_countries %>% 
          select(`Country or Area`, `Sub-region Name`, `Region Name`) %>% 
          mutate(`Country or Area` =  gsub("United Kingdom of Great Britain and Northern Ireland", "United Kingdom", `Country or Area`)) %>% 
          mutate(`Country or Area` =  gsub("United States of America", "USA", `Country or Area`)) %>%
          mutate(`Sub-region Name` =  ifelse(`Sub-region Name` %in% c("Central Asia", "Southern Asia"), "Central and Southern Asia",`Sub-region Name`)) %>%
          mutate(`Sub-region Name` =  ifelse(`Sub-region Name` %in% c("Melanesia", "Micronesia","Polynesia"), "Oceania (others)",`Sub-region Name`)), 
        by.x="country_of_birth", by.y="Country or Area")
  
```



```{r}

table(metadata$`Sub-region Name`)


```


```{r}

cowplot::plot_grid(

metadata %>%
  mutate(age_cat = age_cat %>% forcats::fct_relevel(c("baby","child","teen"), after=0 )) %>%
  filter(!is.na(age_cat)) %>%
  ggplot() + geom_bar(aes(x=age_cat)) + theme_bw(),

metadata %>%
  filter(!is.na(age_cat), !is.na(sex)) %>%
  ggplot() + geom_bar(aes(x=sex)) + theme_bw(), rel_widths = c(2,1))


```



import outliers to filter them out and select the top 30 genus based on read mass.

```{r}
outliers = readLines(con="outliers_samples.txt")

n=30 # n samples per group

metadata %>% 
  filter(!(sample_name %in% outliers)) %>%
  #filter(!(`#SampleID` %in% outliers)) %>%
  merge(country_info, by.x="country_of_birth", by.y="country") %>%
  select(sample_name,age_cat,sex, `Sub-region Name`,country_of_birth) %>% 
  #select(`#SampleID`,age_cat,sex, continent,country_of_birth) %>% 
  filter(sex %in% c("male","female")) %>%
  filter(!is.na(age_cat) & age_cat != "Not provided") %>%
  group_by(age_cat,sex,`Sub-region Name`) %>% 
  sample_n(if(n() < n) n() else n) %>%
  #sample_n(size=30, replace=TRUE) %>%
  #unique() %>% 
  pull(sample_name) -> samples_id_select
  #summarise(n=n())

length(samples_id_select)


metadata %>% 
  filter(sample_name %in% samples_id_select) %>%
  merge(country_info, by.x="country_of_birth", by.y="country") %>%
  select(sample_name,age_cat,sex, `Sub-region Name`,country_of_birth, bmi_cat,gluten,types_of_plants,diet_type  ) %>% 
  filter(sex %in% c("male","female")) %>%
  filter(!is.na(age_cat) & age_cat != "Not provided") %>% #-> metadata_select
  with(., xtabs(~`Sub-region Name`+age_cat, data=.))


metadata %>% 
  filter(sample_name %in% samples_id_select) %>%
  merge(country_info, by.x="country_of_birth", by.y="country") %>%
  select(sample_name,age_cat,sex, `Sub-region Name`,country_of_birth, bmi_cat,gluten,types_of_plants,diet_type  ) %>% 
  filter(sex %in% c("male","female")) %>%
  filter(!is.na(age_cat) & age_cat != "Not provided") -> metadata_select

top_genus_mass = 
  genus_prop %>% 
  select(-outliers) %>%
  tibble::column_to_rownames("taxa") %>% 
  as.matrix %>% 
  apply(1,sum) %>% 
  sort %>% rev %>% head(30)

save(top_genus_mass, file="top_genus_mass.rda")

# genus_prop %>% 
#   select(-outliers) %>%
#   filter(taxa %in% names(top_genus_mass)) %>%
#   summarise_at(-1, sum) %>% t %>% hist()




```



```{r eval=FALSE, include=FALSE}

genus=genus[,colnames(genus) %in% samples_id_select]
source("../notebook/enterotyping.R")

save(fit_genus_list, file="fit_genus_list.rda")


```

load DMM model fit from boostrap
```{r}

#source("enterotypes_bootstrap.R")


# fit_genus_bootstrap = vector("list",5)
# load("../fit_genus_list_1.rda")
# fit_genus_bootstrap[[1]] = fit_genus_list
# load("../fit_genus_list_2.rda")
# fit_genus_bootstrap[[2]] = fit_genus_list
# load("../fit_genus_list_3.rda")
# fit_genus_bootstrap[[3]] = fit_genus_list
# load("../fit_genus_list_4.rda")
# fit_genus_bootstrap[[4]] = fit_genus_list
# load("../fit_genus_list_5.rda")
# fit_genus_bootstrap[[5]] = fit_genus_list

load("../fit_genus_bootstrap.rda")

```


extract BIC and Laplace coefficients from boostrap models  

```{r}
BIC_list = map(fit_genus_bootstrap, function(z) map(z, function(y) sapply(y, function(x){attr(x,"goodnessOfFit")[["BIC"]]}))) 
lplc_list = map(fit_genus_bootstrap, function(z) map(z, function(y) sapply(y, function(x){attr(x,"goodnessOfFit")[["Laplace"]]})))
k = map(fit_genus_bootstrap, function(z) map(z, function(y) sapply(y, function(x){attr(x,"mixture")$Weight %>% length})))[[1]][[1]]


BIC_df = NULL

for(i in 1:5) {
  for(j in 1:5){
    
    BIC_df=cbind(BIC_df,BIC_list[[i]][[j]])
    
  }
}



lplc_df = NULL

for(i in 1:5) {
  for(j in 1:5){
    
    lplc_df=cbind(lplc_df,lplc_list[[i]][[j]])
    
  }
}
```


plot BIC and laplace results
```{r fig.height=6, fig.width=12}

cowplot::plot_grid(

BIC_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:5, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*5))) %>%
  ggplot() + geom_line(aes(x=`Dirichlet component`,y=value,group=variable,color=boot)) +
  ggtitle("BIC optimun") + #xlab("Dirichlet component") +
  theme_bw() + scale_x_continuous(limits = c(1,101), breaks = scales::pretty_breaks(n = 20)),

lplc_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:5, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*5))) %>%
  ggplot() + geom_line(aes(x=`Dirichlet component`,y=value,group=variable,color=boot)) +
  ggtitle("Laplace optimun") + #xlab("Dirichlet component") +
  theme_bw() + scale_x_continuous(limits = c(1,101), breaks = scales::pretty_breaks(n = 20)),



BIC_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:5, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*5))) %>%
  group_by(boot,seed) %>%
  filter(value==min(value)) %>%
  ggplot() + geom_bar(aes(x=`Dirichlet component`)) + 
  ggtitle("BIC optimun") + 
  xlab("Dirichlet components")+
  theme_bw() + scale_x_continuous(limits = c(1,30), breaks = scales::pretty_breaks(n = 30)),


lplc_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k) %>%
  reshape2::melt(id.vars="Dirichlet component") %>% 
  mutate(seed=rep(sort(rep(1:5, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*5))) %>%
  group_by(boot,seed) %>%
  filter(value==min(value)) %>%
  ggplot() + geom_bar(aes(x=`Dirichlet component`)) + 
  ggtitle("Laplace optimun") + 
  xlim(1,30) +  xlab("Dirichlet components")+
  theme_bw() + scale_x_continuous(limits = c(1,30), breaks = scales::pretty_breaks(n = 30)),
labels = ""
) -> figS8ABCD

figS8ABCD

```


BIC and Laplace have a global optimun among differents seeds and bootstraps. All bootstraps found at least one model optimal fit at k=19. Majority of best model based on Laplace were k=19. For BIC, k=19 arrived in second postition. For downstream, analysis we then choose k=19.


```{r}

rbind(lplc_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k, fit="Laplace") %>%
  reshape2::melt(id.vars=c("Dirichlet component","fit")) %>% 
  mutate(seed=rep(sort(rep(1:5, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*5))) ,


BIC_df %>% as.data.frame %>% 
  mutate(`Dirichlet component`=k, fit="BIC") %>%
  reshape2::melt(id.vars=c("Dirichlet component","fit")) %>% 
  mutate(seed=rep(sort(rep(1:5, length(k))),5)) %>%
  mutate(boot=sort(rep(paste0("boot_",1:5), length(k)*5))) ) %>%
  group_by(boot,seed,fit) %>%
  filter(value==min(value)) %>%
  filter(`Dirichlet component`==19) %>%
  arrange(desc(value)) %>%
  ungroup() %>%
  group_by(fit) %>%
  #filter(boot != "boot_1")
  top_n(3,value)

```

For downstream analysis, we choose models among bootstrap where BIC and Laplace showed k=19 as optimum. Among models, we choose the worst one to avoid model overfitting.


```{r fig.height=6, fig.width=12}
boot=1
seed=1
fit_genus_list = fit_genus_bootstrap[[boot]]
best_genus_lplc = 19 
best_genus_lplc_idx = which(sapply(fit_genus_list[[1]], function(x) attr(x,"mixture")$Weight %>% length)==19) # 14th items from list corresponded to k=19 here


enterotypes =

  fit_genus_list[[seed]][[best_genus_lplc_idx]] %>%
  mixture(assign=TRUE) %>% as.data.frame %>% set_colnames(c("Enterotypes_id"))


metadata %>%
  select(sample_name,`Region Name`,country_of_birth,sex,age_cat,age_years,`Sub-region Name`) %>%
  filter(sample_name %in% rownames(enterotypes)) %>%
  with(., xtabs(~`Sub-region Name`+age_cat, data=.))

genus %>%
  select(-outliers) %>%
  tibble::column_to_rownames("taxa") %>% .[names(top_genus_mass),]-> genus_df


#heatmapdmn(t(genus_df[,colnames(genus_df) %in% rownames(enterotypes)]),fit_genus_list[[1]][[1]],fit_genus_list[[1]][[best_genus_lplc]])
```




```{r fig.height=6, fig.width=14}
###################################################
### code chunk number 11: posterior-mean-diff
###################################################
p0 <- fitted(fit_genus_list[[seed]][[1]], scale=TRUE)     # scale by theta
pbest <- fitted(fit_genus_list[[seed]][[best_genus_lplc_idx]], scale=TRUE)
colnames(pbest) <- paste("m", 1:best_genus_lplc, sep="")
(meandiff <- colSums(abs(pbest - as.vector(p0))))
sum(meandiff)


###################################################
### code chunk number 12: table-1
###################################################
diff <- rowSums(abs(pbest - as.vector(p0)))
o <- order(diff, decreasing=TRUE)
cdiff <- cumsum(diff[o]) / sum(diff)
df <- head(cbind(Mean=p0[o], pbest[o,], diff=diff[o], cdiff), 30)
df %>% as.data.frame()

df_gutzilla = df %>% as.data.frame()



df %>% as.data.frame() %>%
  tibble::rownames_to_column("taxa") %>%
  mutate(taxa=factor(taxa, levels=row.names(df))) %>% 
  reshape2::melt(id.vars="taxa") %>%
  filter(!variable %in% c("Mean","diff","cdiff")) %>%
  #group_by(taxa) %>%
  #mutate(total = sum(value)) %>%
  #mutate(value=value/total) %>%
  ggplot() + 
  geom_tile(aes(x=variable,y=taxa,fill=value)) + 
  scale_fill_viridis_c("DMM model alpha") +
  ylab("taxa") +
  xlab("DMM components")
  


```



look at theta per component
```{r}

#fitted(fit_genus_list[[1]][[best_genus_lplc]])

mixturewt(fit_genus_list[[seed]][[1]])

mixturewt(fit_genus_list[[seed]][[best_genus_lplc_idx]])

mixturewt(fit_genus_list[[seed]][[best_genus_lplc_idx]]) %>% 
  tibble::rowid_to_column("k") %>% 
  ggplot() + 
  geom_bar(aes(x=k,y=theta),stat="identity") + 
  geom_hline(yintercept = mixturewt(fit_genus_list[[seed]][[1]])$theta, col="red") +
  cowplot::theme_cowplot() +
  ylab("theta") +
  xlab("DMM community types") -> figS8E

figS8E

```



```{r}

metadata %>%
  merge(enterotypes, by.x="sample_name", by.y="row.names", all.y="TRUE") %>%
  group_by(`Sub-region Name`) %>%
  mutate(total=n()) %>%
  group_by(`Sub-region Name`,Enterotypes_id,total) %>%
  summarise(prop=n()) %>%
  mutate(prop=prop/total) %>%
  ggplot() + geom_tile(aes(x=Enterotypes_id %>% as.character(),y=`Sub-region Name`,fill=prop)) + scale_fill_viridis_c()


metadata %>%
  merge(enterotypes, by.x="sample_name", by.y="row.names", all.y="TRUE") %>%
  filter(age_cat != "baby") %>%
  group_by(age_cat) %>%
  mutate(total=n()) %>%
  group_by(age_cat,Enterotypes_id,total) %>%
  summarise(prop=n()) %>%
  mutate(prop=prop/total) %>%
  ungroup() %>%
  group_by(Enterotypes_id) %>%
  mutate(prop2=sum(prop)) %>%
  mutate(prop=prop/prop2) %>%
  ggplot() + geom_tile(aes(x=Enterotypes_id %>% as.character(),y=age_cat,fill=prop)) + scale_fill_viridis_c()


metadata %>%
  merge(enterotypes, by.x="sample_name", by.y="row.names", all.y="TRUE") %>%
  filter(age_cat != "baby") %>%
  filter(!is.na(bmi_cat), bmi_cat != "Not provided") %>%
  group_by(bmi_cat) %>%
  mutate(total=n()) %>%
  group_by(bmi_cat,Enterotypes_id,total) %>%
  summarise(prop=n()) %>%
  mutate(prop=prop/total) %>%
  ungroup() %>%
  group_by(bmi_cat) %>%
  mutate(prop2=sum(prop)) %>%
  mutate(prop=prop/prop2) %>%
  ggplot() + geom_tile(aes(x=Enterotypes_id %>% as.character(),y=bmi_cat,fill=prop)) + scale_fill_viridis_c()


metadata %>%
  merge(enterotypes, by.x="sample_name", by.y="row.names", all.y="TRUE") %>%
  filter(age_cat != "baby") %>%
  filter(!is.na(gluten), gluten != "Not provided") %>%
  group_by(gluten) %>%
  mutate(total=n()) %>%
  group_by(gluten,Enterotypes_id,total) %>%
  summarise(prop=n()) %>%
  mutate(prop=prop/total) %>%
  ungroup() %>%
  group_by(gluten) %>%
  mutate(prop2=sum(prop)) %>%
  mutate(prop=prop/prop2) %>%
  ggplot() + geom_tile(aes(x=Enterotypes_id %>% as.character(),y=gluten,fill=prop)) + scale_fill_viridis_c()

metadata %>%
  merge(enterotypes, by.x="sample_name", by.y="row.names", all.y="TRUE") %>%
  filter(age_cat != "baby") %>%
  filter(!is.na(types_of_plants), types_of_plants != "Not provided") %>%
  group_by(types_of_plants) %>%
  mutate(total=n()) %>%
  group_by(types_of_plants,Enterotypes_id,total) %>%
  summarise(prop=n()) %>%
  mutate(prop=prop/total) %>%
  ungroup() %>%
  group_by(types_of_plants) %>%
  mutate(prop2=sum(prop)) %>%
  mutate(prop=prop/prop2) %>%
  ggplot() + geom_tile(aes(x=Enterotypes_id %>% as.character(),y=types_of_plants,fill=prop)) + scale_fill_viridis_c()

metadata %>%
  merge(enterotypes, by.x="sample_name", by.y="row.names", all.y="TRUE") %>%
  filter(age_cat != "baby") %>%
  filter(!is.na(diet_type), diet_type != "Not provided") %>%
  group_by(diet_type) %>%
  mutate(total=n()) %>%
  group_by(diet_type,Enterotypes_id,total) %>%
  summarise(prop=n()) %>%
  mutate(prop=prop/total) %>%
  ungroup() %>%
  group_by(diet_type) %>%
  mutate(prop2=sum(prop)) %>%
  mutate(prop=prop/prop2) %>%
  ggplot() + geom_tile(aes(x=Enterotypes_id %>% as.character(),y=diet_type,fill=prop)) + scale_fill_viridis_c()

```


to do: compute enterotype proportion for all categories and do a PCA analysis with minimal spanning tree


```{r}

metadata %>%
  merge(enterotypes, by.x="sample_name", by.y="row.names", all.y="TRUE") %>%
  select(sample_name,age_cat,sex,`Sub-region Name`,bmi_cat, Enterotypes_id) %>%
  reshape2::melt(id.vars=c("sample_name","Enterotypes_id")) %>%
  filter(!is.na(value), value != "Not provided") %>%
  select(-variable, -`sample_name`) %>%
  group_by(Enterotypes_id,value) %>%
  filter(!is.na(value)) %>%
  #mutate(total=n()) %>%
  #group_by(Enterotypes_id,value,total) %>%
  #summarise(prop=n()) %>%
  reshape2::dcast(Enterotypes_id~value) %>%
  mutate_at(-1,~./(sum(.))) -> et_repartition

et_repartition %>%
  tibble::column_to_rownames("Enterotypes_id") %>%
  apply(1,function(x) x/sum(x)) %>%
  t %>%
  as.data.frame %>%
  ade4::dudi.pca(.,scan=F, nf=3) -> et_metadata_pca


ade4::scatter(et_metadata_pca)

ade4::s.label(et_metadata_pca$li, yax=2)

```


```{r fig.height=5, fig.width=12}

df %>% as.data.frame() %>%
  select(-diff,-cdiff,-Mean) %>%
  mutate_at(1:19,~log10(.)) %>%
  t %>%
  as.data.frame %>%
  ade4::dudi.pca(scannf = F,nf=3) %>% .$li %>% ade4::s.label()




df %>% as.data.frame() %>%
  select(-diff,-cdiff,-Mean) %>%
  tibble::rownames_to_column("taxa") %>%
  #mutate(taxa2=taxa) %>%
  #mutate(taxa2=ifelse())
    as_tibble %>%
  .[c(1:3,7,11),] %>%
  reshape2::melt(id.vars="taxa") %>%
  group_by(variable) %>%
  mutate(ymax=cumsum(value)) %>%
  mutate(ymin=c(0, head(ymax, n=-1))) %>%
  #ggplot() + geom_bar(aes(y=value,x=variable,fill=taxa),stat="identity") 
  ggplot(aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=taxa %>% strsplit(x=.,split=";") %>% sapply(function(x)x[6])  )) + 
  geom_rect() +      
  coord_polar(theta="y") + 
  facet_wrap(~variable) + 
  xlim(c(2, 4)) + theme_void() +
  scale_fill_brewer("",type="qual")

ggsave("agp_microbiota_types.pdf")

```

reclassify all dataset
```{r}

# fit_genus_list[[1]][[best_genus_lplc]] %>%
#   mixture(assign=TRUE) %>% as.data.frame %>% set_colnames(c("Enterotypes_id"))

genus_select=rownames(fitted(fit_genus_bootstrap[[1]][[1]][[best_genus_lplc]]))
genus %>%
      select(-outliers) %>%
      tibble::column_to_rownames("taxa") %>% .[genus_select,]-> genus_df

genus_df %>% colnames() %>% data.frame(sample_name=.) -> enterotypes_clusters_df


 k=1
for(i in 1:5) {
  
  for(j in 1:5){
    

   
    
    predict(fit_genus_bootstrap[[i]][[j]][[best_genus_lplc]], t(genus_df), assign = TRUE) %>%
      reshape2::melt() %>%
      group_by(Var1) %>%
      filter(value==max(value)) %>%
      dplyr::rename(`#SampleID`="Var1", Enterotypes_id="Var2") %>%
      select(`#SampleID`,Enterotypes_id) -> enterotypes_prediction_max
    
    colnames(enterotypes_prediction_max) = c("sample_name",paste0("Enterotypes_id_",k))
    
    enterotypes_clusters_df=
    enterotypes_clusters_df %>%
      merge(enterotypes_prediction_max, by="sample_name")
    
    cat(k,"\n")
    k=k+1
   
    
  }
  
  
}

 enterotypes_clusters_dist = matrix(nr=25,nc=25)  

 for(i in 1:25) {
     for (j in 1:25) {
     
     enterotypes_clusters_dist[i,j] = mclust::adjustedRandIndex(enterotypes_clusters_df[-1][,i],enterotypes_clusters_df[-1][,j])
     
     
   }
 }
 
 

```

### classify data with the choosen model


```{r}



predict(fit_genus_list[[1]][[14]], t(genus_df), assign = TRUE) %>%
  reshape2::melt() %>%
  group_by(Var1) %>%
  filter(value==max(value)) %>%
  dplyr::rename(`sample_name`="Var1", Enterotypes_id="Var2") %>%
  mutate(set=ifelse(sample_name %in% row.names(attr(fit_genus_list[[1]][[14]],"group")), "train", "predicted")) -> enterotypes_prediction


enterotypes_prediction %>%
  ggplot() + geom_histogram(aes(x=value, fill=value>0.80), bins = 50) + 
  #facet_wrap(~set) + 
  scale_x_continuous(limits = c(0,1.05), labels = scales::percent)

enterotypes_prediction %>%
  ggplot() + 
  geom_histogram(aes(x=value), bins = 50) + 
  #facet_wrap(~set) + 
  scale_x_continuous(limits = c(0,1.05), labels = scales::percent) +
  cowplot::theme_cowplot() -> figS8F


```

### export enterotypes with outliers

```{r}

enterotypes_prediction %>%
  as.data.frame %>%
  rbind(data.frame(sample_name = outliers,Enterotypes_id=NA,value=NA,set="outliers", stringsAsFactors = FALSE)) %>%
  write.csv2(file="enterotypes_prediction_outliers.csv")


```






## Enterotypes vs Age and Richness


### Age


```{r}

metadata %>%
  merge(enterotypes, by.x="sample_name", by.y="row.names", all.y="TRUE") %>%
  select(Enterotypes_id, age_years) %>% ggplot() + geom_boxplot(aes(x=Enterotypes_id%>% as.character, y=age_years))
  group_by(Enterotypes_id) %>%
  summarise(age_years=mode(age_years))
  

```



## richness


```{r}

shannon_path = system.file("data-raw/qiime/generated-files-20190512/alpha/shannon.qza", package = "gutzilla")
shannon = qiime2R::read_qza(shannon_path)$data %>% as.data.frame

merge(enterotypes_prediction,shannon,by.x="sample_name", by.y="row.names")  -> enterotypes_alpha_div
  
enterotypes_alpha_div %>% 
  ggplot()   + 
  geom_boxplot(aes(x=paste0("M",Enterotypes_id %>% as.character),y=shannon, fill=set)) +
  xlab("DMM community types") + ylab("alpha-diversity (shannon)") +
  cowplot::theme_cowplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) -> figS8G
figS8G



```

## figure supp 8


```{r fig.height=10, fig.width=16}

(figS8ABCD/(figS8E+figS8F+figS8G + plot_layout(widths = c(1,1,1.5)))) + plot_layout(heights = c(2,1)) + plot_annotation(tag_levels = 'A') -> figS8

figS8
ggsave("figS8.pdf")



```



## check fibres


```{r}

fibre_path = system.file("data-raw/vioscreen/FICHIER_JULIEN_FIBER_HOSTSUBJECTID_2.xlsx", package = "agp")

readxl::read_xlsx(fibre_path) %>%
  select(-FIBER...6) %>%
  dplyr::rename(FIBER="FIBER...4") -> fibre_df
  

colnames(fibre_df) = colnames(fibre_df) %>% stringr::str_to_lower()

```

n=1275 with fiber and enterotypes data

```{r}

fibre_df %>%
  merge(enterotypes_alpha_div, by="sample_name") %>% 
  ggplot() + 
  geom_boxplot(aes(y=shannon,x=paste0("Q",quartile_fiber),fill=paste0("Q",quartile_fiber))) +
  xlab("fiber\nquartiles") + ylab("alpha-diversity (shannon)") +
  scale_fill_brewer("fiber\nquartiles",type="div",palette = 2)


fibre_df %>%
  merge(enterotypes_alpha_div, by="sample_name") %>% 
  ggplot() + 
  geom_boxplot(aes(y=shannon,x=paste0("m",Enterotypes_id %>% as.character),fill=paste0("Q",quartile_fiber))) +
  xlab("Microbiota DMM based clusters") + ylab("alpha-diversity (shannon)") +
  scale_fill_brewer("fiber\nquartiles",type="div",palette = 2)


fibre_df %>%
  merge(enterotypes_alpha_div, by="sample_name") %>% 
  group_by(Enterotypes_id) %>%
  do(with(.,kruskal.test(.$shannon,.$quartile_fiber, method="spearman") %>% broom::tidy() ))

fibre_df %>%
  merge(enterotypes_alpha_div, by="sample_name") %>% 
  group_by(Enterotypes_id) %>%
  summarise(n=n())

fibre_df %>%
  merge(enterotypes_alpha_div, by="sample_name") %>% 
  filter(Enterotypes_id %in% c(2,5,6,7,13)) %>%
  ggplot() + 
  geom_boxplot(aes(y=shannon,x=paste0("m",Enterotypes_id %>% as.character),fill=paste0("Q",quartile_fiber))) +
  xlab("Microbiota DMM based clusters") + ylab("alpha-diversity (shannon)") +
  scale_fill_brewer("fiber\nquartiles",type="div",palette = 2)




```

## check type of plant


```{r fig.height=5, fig.width=8}



metadata %>%
  select(sample_name,types_of_plants) %>%
   mutate(types_of_plants=types_of_plants%>% forcats::fct_rev() %>% forcats::fct_relevel("More than 30", after=Inf) %>% forcats::fct_relevel("21 to 30", after=3)) %>%
  na.omit() %>%
  merge(enterotypes_alpha_div, by="sample_name") %>% 
  ggplot() + 
  geom_boxplot(aes(y=shannon,x=types_of_plants,fill=types_of_plants)) +
  xlab("nb type of plants per week") + ylab("alpha-diversity (shannon)") +
  scale_fill_brewer("Type of plants",type="div",palette = 2) +
  xlab("Type of plants") + ylab("alpha-diversity (shannon)") + ggtitle("all pop")

metadata %>%
  select(sample_name,types_of_plants) %>%
  merge(fibre_df, by="sample_name") %>%
  mutate(types_of_plants=types_of_plants%>% forcats::fct_rev() %>% forcats::fct_relevel("More than 30", after=Inf) %>% forcats::fct_relevel("21 to 30", after=3)) %>%
  na.omit() %>%
  merge(enterotypes_alpha_div, by="sample_name") %>% 
  ggplot() + 
  geom_boxplot(aes(y=shannon,x=types_of_plants,fill=types_of_plants)) +
  xlab("nb type of plants per week") + ylab("alpha-diversity (shannon)") +
  scale_fill_brewer("Type of plants",type="div",palette = 2) +
  xlab("Type of plants ") + ylab("alpha-diversity (shannon)") + ggtitle("with vioscreened only")



metadata %>%
  select(sample_name,types_of_plants) %>%
   mutate(types_of_plants=types_of_plants%>% forcats::fct_rev() %>% forcats::fct_relevel("More than 30", after=Inf) %>% forcats::fct_relevel("21 to 30", after=3)) %>%
  na.omit() %>%
  merge(enterotypes_alpha_div, by="sample_name") %>% 
  ggplot() + 
  geom_boxplot(aes(y=shannon,x=paste0("m",Enterotypes_id %>% as.character),fill=types_of_plants)) +
  xlab("Microbiota DMM based clusters") + ylab("alpha-diversity (shannon)") +
  scale_fill_brewer("Type of\nplant",type="div",palette = 2) + ggtitle("all pop")



metadata %>%
  select(sample_name,types_of_plants) %>%
  merge(fibre_df, by="sample_name") %>%
   mutate(types_of_plants=types_of_plants%>% forcats::fct_rev() %>% forcats::fct_relevel("More than 30", after=Inf) %>% forcats::fct_relevel("21 to 30", after=3)) %>%
  na.omit() %>%
  merge(enterotypes_alpha_div, by="sample_name") %>% 
  ggplot() + 
  geom_boxplot(aes(y=shannon,x=paste0("m",Enterotypes_id %>% as.character),fill=types_of_plants)) +
  xlab("Microbiota DMM based clusters") + ylab("alpha-diversity (shannon)") +
  scale_fill_brewer("Type of\nplant",type="div",palette = 2) + ggtitle("with vioscreened only")





```


```{r}

metadata %>%
  select(sample_name,types_of_plants) %>%
  merge(fibre_df, by="sample_name") %>%
   mutate(types_of_plants=types_of_plants%>% forcats::fct_rev() %>% forcats::fct_relevel("More than 30", after=Inf) %>% forcats::fct_relevel("21 to 30", after=3)) %>%
  na.omit() %>%
  merge(enterotypes_alpha_div, by="sample_name") %>%
  filter(Enterotypes_id %in% c(1,2,5,7,13,15)) %>%
  ggplot() + geom_point(aes(x=fiber,y=shannon,col=types_of_plants)) + 
  scale_color_brewer("Type of\nplant",type="div",palette = 2) + 
  facet_wrap(~paste0("m",Enterotypes_id), scales = "free_y") + theme_dark() +
  ggtitle("example of enterotypes")




```




## rescale enterotypes using world demographic data

we will use UNSD subregion to rescale enterotypes proportion observed

```{r}

data("UNSD_pop_subregion")

UNSD_pop_subregion %>%
          mutate(Subregion = ifelse(Subregion %in% c("Melanesia","Micronesia","Polynesia"), "Oceania (others)",Subregion)) %>%
          mutate(Subregion =  ifelse(Subregion %in% c("Central Asia", "Southern Asia"), "Central and Southern Asia",Subregion)) %>%
          mutate(Subregion =  ifelse(Subregion %in% c("South-Eastern Asia"), "South-eastern Asia",Subregion)) %>%
  ggplot() + geom_bar(aes(y=n_pop_2020, x=Subregion), stat="identity") +
  theme(axis.text.x = element_text(angle = -90, vjust=0.5))



metadata %>%
  merge(enterotypes_prediction, by.x="sample_name", all.y="TRUE") %>%
  select(Enterotypes_id,`Sub-region Name`) %>%
  na.omit %>%
  group_by(`Sub-region Name`) %>%
  summarise(total=n()) %>%
  ggplot() + geom_bar(aes(y=total, x=`Sub-region Name`), stat="identity") +
  theme(axis.text.x = element_text(angle = -90, vjust=0.5))

metadata %>%
  merge(enterotypes_prediction, by.x="sample_name",  all.y="TRUE") %>%
  select(`Sub-region Name`) %>%
  na.omit %>%
  group_by(`Sub-region Name`) %>%
  summarise(total=n()) %>%
  merge(UNSD_pop_subregion %>%
          mutate(Subregion = ifelse(Subregion %in% c("Melanesia","Micronesia","Polynesia"), "Oceania (others)",Subregion)) %>%
          mutate(Subregion =  ifelse(Subregion %in% c("Central Asia", "Southern Asia"), "Central and Southern Asia",Subregion)) %>%
          mutate(Subregion =  ifelse(Subregion %in% c("South-Eastern Asia"), "South-eastern Asia",Subregion)) %>%
                   group_by(Subregion) %>%
                   summarise(n_pop_2020 = sum(n_pop_2020)),
  by.x="Sub-region Name", by.y="Subregion", all.x=TRUE) %>%
  mutate_if(is.numeric,function(x) x/sum(x) ) %>%
  dplyr::rename(n_pop_agp_birth="total") -> region_n_pop_agp_unsd

  region_n_pop_agp_unsd %>%
  reshape2::melt(id.vars="Sub-region Name") %>%
  ggplot() + 
  geom_bar(aes(x=`Sub-region Name`, y=value, fill=variable), stat="identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = -90, vjust=0.5)) 

  
  
  



```

```{r}

metadata %>%
  merge(enterotypes_prediction, by.x="sample_name",  all.y="TRUE") %>%
  select(`Sub-region Name`,Enterotypes_id,set) %>%
  ggplot() + geom_bar(aes(x=Enterotypes_id,fill=set)) 

metadata %>%
  merge(enterotypes_prediction, by.x="sample_name",  all.y="TRUE") %>%
  select(`Sub-region Name`,Enterotypes_id,set) %>%
  na.omit %>%
  group_by(`Sub-region Name`,Enterotypes_id) %>%
  summarise(n_per_et=n()) %>%
  ungroup() %>%
  group_by(`Sub-region Name`) %>%
  mutate(n_per_region=sum(n_per_et)) %>%
  ungroup() %>%
  #mutate(n=n %>% (function(x) x / sum(x)) ) %>%
  merge(UNSD_pop_subregion %>%
          filter(!(Subregion == "Latin America and the Caribbean" & Type =="Region")) %>%
          mutate(Subregion = ifelse(Subregion %in% "Oceania (Excluding Australia and New Zealand)", "Oceania (others)",Subregion)) %>%
          #mutate(Subregion =  ifelse(Subregion %in% c("Central Asia", "Southern Asia"), "Central and Southern Asia",Subregion)) %>%
          mutate(Subregion =  ifelse(Subregion %in% c("South-Eastern Asia"), "South-eastern Asia",Subregion)) ,
                   
                   #summarise(n_pop_2020 = sum(n_pop_2020)),
  by.x="Sub-region Name", by.y="Subregion", all.x=TRUE) %>%
  mutate(n_estimated_per_pop = (n_per_et * n_pop_2020)/n_per_region) %>%
  select(-n_per_region,-Type,-n_pop_2020) %>%
  mutate(p_per_et=n_per_et/sum(n_per_et), p_estimated_per_pop=n_estimated_per_pop/sum(n_estimated_per_pop))  -> enterotypes_by_UNSD_regions

write.csv2(enterotypes_by_UNSD_regions, file="enterotypes_by_UNSD_regions.csv")
  
  
enterotypes_by_UNSD_regions %>%
  group_by(Enterotypes_id) %>%
  summarise_if(is.numeric,sum) %>%
  select(Enterotypes_id,p_per_et,p_estimated_per_pop) %>%
  reshape2::melt(id.vars="Enterotypes_id") %>%
  ggplot() + geom_bar(aes(x=Enterotypes_id,y=value,fill=variable), stat="identity", position = "dodge")


```




