---
title: "GUTZILLA visual examples"
output: html_notebook
---


```{r}
library(biomformat)
library(tidyverse)
library(qiime2R)
devtools::load_all()

```

load metadata
```{r}

metadata = readr::read_tsv(system.file("data-raw/qiime/generated-files-20190512/ag.txt", package = "agp"))

country_info = read_tsv(system.file(package="agp", "data-raw/countryInfo.txt"), comment = "#", col_names = FALSE, na = "") %>%
  select(X5,X9) %>%
  dplyr::rename(country="X5", continent="X9" )

```


load genus abundance
```{r}

genus_path = system.file("data-raw/qiime/generated-files-20190512/taxa/genus.qza", package = "agp")

genus = qiime2R::read_qza(genus_path)$data %>% as.data.frame %>% tibble::rownames_to_column("taxa")  %>% as_tibble()


```


normalize genus abundance
```{r}
genus %>%
  #select(1:10) %>%
  mutate_at(-1, ~./sum(.)) -> genus_prop


# genus_prop %>%
#   tibble::column_to_rownames("taxa") %>%
#   BiotypeR::noise.removal(percent=1) %>%
#   row.names() -> genus_select


# genus_prop %>%
#   tibble::rownames_to_column("taxa") %>%
#   filter(taxa %in% genus_select) %>%
#   reshape2::melt(id.vars="taxa")
```

select the top 10 genus for adult and child

> I remove unclassified genera

```{r}

top_genus_mass = 
  genus_prop %>% 
  tibble::column_to_rownames("taxa") %>% 
  as.matrix %>% 
  apply(1,sum) %>% 
  sort %>% rev %>% head(50)


genus_prop %>%
  filter(taxa %in% c(top_genus_mass %>% names())) %>%
  dplyr::rename(Genus = "taxa") %>%
  #select(1:20) %>%
  #tidyr::separate(taxa, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  #select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  #mutate(Genus = Genus %>% gsub("g__","",.)) %>%
  reshape2::melt(id.vars="Genus") -> top_genus_prop_value_50

metadata %>%
  select("#SampleID","age_cat","age_corrected","sex","country_of_birth","country") %>%
  merge(top_genus_prop_value_50, by.x="#SampleID", by.y="variable") -> top_genus_prop_value_metadata

#top_genus_prop_value_metadata$age_cat %>% table 

top_genus_prop_value_metadata %>%
  filter(!is.na(age_cat), age_cat!="Not provided") %>%
  mutate(Adult=ifelse(age_cat %in% c("baby","teen","child"),"Child","Adult")) %>%
  filter(Genus != "__") %>% 
  group_by(Adult,Genus) %>% 
  summarise(mass=sum(value)) %>%
  reshape2::dcast(Genus~Adult) %>%
  mutate_at(2:3, ~./(sum(.))) %>%
  mutate(Adult=rank(-Adult),Child=rank(-Child)) %>%
  filter(Adult <= 10 | Child <= 10) %>%
  arrange(Adult) %>%
  filter(!(Genus %in% c("k__Bacteria;p__Firmicutes;c__Clostridia;o__Clostridiales;f__;g__","k__Bacteria;p__Firmicutes;c__Clostridia;o__Clostridiales;f__Ruminococcaceae;g__"))) %>%
  pull(Genus) -> top_10_genus

writeLines(top_10_genus, con = "genus_top_10.txt")

 
```


vizualise top 10 genus median abundance

```{r fig.height=3, fig.width=6}
# genus_prop %>%
#   filter(taxa %in% top_10_genus) %>%
#   #select(1:20) %>%
#   tidyr::separate(taxa, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
#   select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
#   mutate(Genus = Genus %>% gsub("g__","",.)) %>%
#   reshape2::melt(id.vars="Genus") %>%
#   ggplot() + geom_boxplot(aes(x=Genus,y=value+0.0001)) + 
#   coord_flip() + scale_y_log10(labels = scales::percent) + 
#   ylab("Relative abundance") + xlab("Microbiota genus")
  

genus_prop %>%
  filter(taxa %in% top_10_genus) %>%
  #select(1:20) %>%
  tidyr::separate(taxa, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  mutate(Genus = Genus %>% gsub("g__","",.)) %>%
  reshape2::melt(id.vars="Genus") %>%
  group_by(Genus) %>%
  summarise(med=median(value)) %>%
  ggplot() + geom_bar(aes(x=Genus,y=med), stat="identity") + 
  coord_flip() + #scale_y_log10(labels = scales::percent) + 
  ylab("Relative abundance") + xlab("Microbiota genus")
```

visualise top 10 genera throught age and sex

```{r fig.height=4, fig.width=10}
  
top_genus_prop_value_metadata %>%
  filter(Genus %in% top_10_genus) %>%
  tidyr::separate(Genus, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  mutate(Genus = Genus %>% gsub("g__","",.)) %>%
  filter(sex %in% c("male","female")) %>%
  group_by(age_cat,Genus,sex) %>%
  summarise(value=median(value)) %>%
  filter(!is.na(age_cat) & age_cat != "Not provided") %>%
  ungroup() %>%
  mutate(age_cat=age_cat %>% forcats::fct_relevel("baby","child","teen",after=0)) %>%
  ggplot + geom_point(aes(y=value+0.0001,x=age_cat,color=sex), stat="identity") + 
  geom_line(aes(y=value+0.0001,x=age_cat,group=paste0(Genus,sex), color=sex)) + 
  facet_wrap(~Genus, scale="free",nrow=2) + 
  coord_flip() + scale_y_log10(labels = scales::percent) + xlab("") + ylab("")






```

top 10 genus trought age and continent

```{r fig.height=8, fig.width=12}

library(ggridges)

metadata %>%
  select(1,country_of_birth) %>%
  merge(country_info, by.x="country_of_birth", by.y="country") %>%
  group_by(continent) %>%
  summarise(n=n())

top_genus_prop_value_metadata %>%
  filter(Genus %in% top_10_genus) %>%
  merge(country_info, by.x="country_of_birth", by.y="country") %>%
  tidyr::separate(Genus, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  mutate(Genus = Genus %>% gsub("g__","",.)) %>%
  #filter(sex %in% c("male","female")) %>%
  #group_by(age_cat,Genus) %>%
  #summarise(value=median(value)) %>%
  filter(!is.na(age_cat) & age_cat != "Not provided") %>%
  #filter(country_of_birth %in% c("United States","United Kingdom")) %>%
  #ungroup() %>%
  mutate(age_cat=age_cat %>% forcats::fct_relevel("baby","child","teen",after=0)) %>%
  ggplot + 
  geom_density_ridges_gradient(aes(x=log10(value + 10^-4), y=age_cat, fill=..x..)) + 
  scale_fill_viridis_c() +
  #geom_point(aes(y=value+0.0001,x=age_cat,color=sex), stat="identity") + 
  #geom_line(aes(y=value+0.0001,x=age_cat,group=paste0(Genus,sex), color=sex)) + 
  facet_grid(continent~Genus, scale="free_x") + 
  xlab("") + ylab("")


top_genus_prop_value_metadata %>%
  filter(Genus %in% top_10_genus) %>%
  merge(country_info, by.x="country_of_birth", by.y="country") %>%
  tidyr::separate(Genus, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  mutate(Genus = Genus %>% gsub("g__","",.)) %>%
  #filter(sex %in% c("male","female")) %>%
  #group_by(age_cat,Genus) %>%
  #summarise(value=median(value)) %>%
  #filter(!is.na(age_cat) & age_cat != "Not provided") %>%
  #filter(country_of_birth %in% c("United States","United Kingdom")) %>%
  #ungroup() %>%
  #mutate(age_cat=age_cat %>% forcats::fct_relevel("baby","child","teen",after=0)) %>%
  ggplot + 
  geom_density_ridges_gradient(aes(x=log10(value + 10^-3),y="y", fill=..x..)) + 
  scale_fill_viridis_c() +
  #geom_point(aes(y=value+0.0001,x=age_cat,color=sex), stat="identity") + 
  #geom_line(aes(y=value+0.0001,x=age_cat,group=paste0(Genus,sex), color=sex)) + 
  facet_grid(continent~Genus, scale="free_x") + 
  xlab("") + ylab("") + xlim(-4,0)


```

evaluate top 10 genus read mass per age and continent

```{r}



top_genus_prop_value_metadata %>%
  filter(Genus %in% top_10_genus) %>%
  merge(country_info, by.x="country_of_birth", by.y="country") %>%
  tidyr::separate(Genus, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  mutate(Genus = Genus %>% gsub("g__","",.)) %>%
  filter(!is.na(age_cat) & age_cat != "Not provided") %>%
  mutate(age_cat=age_cat %>% forcats::fct_relevel("baby","child","teen",after=0)) %>%
  group_by(`#SampleID`,country_of_birth,age_cat) %>%
  summarise(mass=sum(value)) %>%
  group_by(age_cat) %>%
  summarise(mass=median(mass))


top_genus_prop_value_metadata %>%
  filter(Genus %in% top_10_genus) %>%
  merge(country_info, by.x="country_of_birth", by.y="country") %>%
  tidyr::separate(Genus, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  mutate(Genus = Genus %>% gsub("g__","",.)) %>%
  filter(!is.na(age_cat) & age_cat != "Not provided") %>%
  mutate(age_cat=age_cat %>% forcats::fct_relevel("baby","child","teen",after=0)) %>%
  filter(age_cat %in% c("20s","30s","40s","50s")) %>%
  group_by(`#SampleID`,country_of_birth,continent, age_cat) %>%
  summarise(mass=sum(value)) %>%
  group_by(continent) %>%
  summarise(mass=median(mass))




```

pull out individual with low top 10 read mass
```{r}


top_genus_prop_value_metadata %>%
  filter(!is.na(age_cat), age_cat!="Not provided") %>%
  mutate(Adult=ifelse(age_cat %in% c("baby","teen","child"),"Child","Adult")) %>%
  #filter(Genus != "__") %>% 
  group_by(Adult,Genus) %>% 
  summarise(mass=sum(value)) %>%
  reshape2::dcast(Genus~Adult) %>%
  mutate_at(2:3, ~./(sum(.))) %>%
  mutate(Adult=rank(-Adult),Child=rank(-Child)) %>%
  filter(Adult <= 10 | Child <= 10) %>%
  arrange(Adult) %>%
  #filter(!(Genus %in% c("k__Bacteria;p__Firmicutes;c__Clostridia;o__Clostridiales;f__;g__","k__Bacteria;p__Firmicutes;c__Clostridia;o__Clostridiales;f__Ruminococcaceae;g__"))) %>%
  pull(Genus) -> top_12_genus



top_genus_prop_value_metadata %>%
  filter(Genus %in% top_12_genus) %>%
  merge(country_info, by.x="country_of_birth", by.y="country") %>%
  tidyr::separate(Genus, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  mutate(Genus = Genus %>% gsub("g__","",.)) %>%
  filter(!is.na(age_cat) & age_cat != "Not provided") %>%
  mutate(age_cat=age_cat %>% forcats::fct_relevel("baby","child","teen",after=0)) %>%
  group_by(`#SampleID`,country_of_birth,age_cat) %>%
  summarise(mass=sum(value)) %>% ggplot() + geom_histogram(aes(x=mass))
  filter(mass<0.25) %>% pull(`#SampleID`) -> microbiota_outliers
  

metadata %>%
  #filter(`#SampleID` %in% microbiota_outliers) %>%
  pull(subset_antibiotic_history) %>% table %>% prop.table

#metadata 

top_genus_prop_value_metadata %>%
  filter(Genus %in% top_12_genus) %>%
  merge(country_info, by.x="country_of_birth", by.y="country") %>%
  tidyr::separate(Genus, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  mutate(Genus = Genus %>% gsub("g__","",.)) %>%
  filter(!is.na(age_cat) & age_cat != "Not provided") %>%
  mutate(age_cat=age_cat %>% forcats::fct_relevel("baby","child","teen",after=0)) %>%
  group_by(`#SampleID`) %>%
  summarise(mass=sum(value)) %>%
  select(1,mass) -> sample_read_mass



```



import pcoa coordinates (unweighted unifrac)
```{r}

url="data-raw/qiime/generated-files-20190512/beta/pcoa/d3609a45-53a8-40e1-b731-812b8127903b/data/ordination.txt" #unweighted
#url="data-raw/qiime/generated-files-20190512/beta/pcoa/d3048346-f38b-4797-9fb3-bd4166d5028d/data/ordination.txt" #weighted

pcoa=
readLines(system.file(url, package="agp"))[-c(1:10)] %>% 
  as.matrix %>% as.data.frame %>% tidyr::separate("V1", into=as.character(1:11), sep="\t") %>%
  select(1:3) %>%
  dplyr::rename(sampleID="1",pcoa1="2",pcoa2="3")


top_genus_prop_value_metadata %>%
  filter(Genus %in% top_10_genus) %>%
  tidyr::separate(Genus, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  mutate(Genus = Genus %>% gsub("g__","",.)) %>% 
  merge(pcoa %>% mutate(pcoa1=as.numeric(pcoa1),pcoa2=as.numeric(pcoa2)) , by.x="#SampleID", by.y="sampleID") %>%
  merge( 
          metadata %>%
  select("#SampleID","age_cat","age_corrected","sex","country_of_birth","country") %>% unique,
  
  by.x="#SampleID", by.y="#SampleID"
  
          )  -> test
```


top 10 genus vs PCoA
```{r}
test %>%
  ggplot + geom_hex(aes(value,pcoa1)) + scale_color_viridis_c()  + facet_wrap(~Genus)

test %>%
  ggplot + geom_hex(aes(value,pcoa2)) + scale_color_viridis_c()  + facet_wrap(~Genus)
```


```{r fig.height=4, fig.width=10}
test %>%
  ggplot + geom_point(aes(pcoa1, pcoa2, color=log10(value+0.0001) ), size=1) +  facet_wrap(~Genus, nrow=2) + scale_color_viridis_c() 

# 
# test %>%
#   ggplot + geom_point(aes(pcoa1, pcoa2, color=log10(value+0.0001) ), size=1) +  facet_grid(age_cat.x~Genus) + scale_color_viridis_c()
```


PCoA bins vs age average
```{r}
test %>%
  filter(Genus=="Bifidobacterium") %>%
  ggplot(aes(pcoa1, pcoa2, z=as.numeric(age_corrected.x))) + 
  stat_summary_hex(bins = 50) + 
    scale_fill_viridis_c("Age mean") + theme_void()
```


```{r}
enterotypes=read.table("enterotypes_12.txt", sep="\t")

test %>%
  filter(Genus=="Bifidobacterium") %>%
  merge(enterotypes, by.x="#SampleID", by.y="row.names") %>%
  ggplot(aes(pcoa1, pcoa2, color=as.character(Enterotypes_id) )) + 
  geom_point() +
  facet_wrap(~Enterotypes_id) +
  theme_dark()


```



```{r}
test %>%
  filter(Genus=="Bifidobacterium") %>%
  mutate(outliers = ifelse(`#SampleID` %in% microbiota_outliers,"outliers","ok")) %>%
  merge(sample_read_mass, by="#SampleID") %>%
  ggplot(aes(pcoa1, pcoa2)) + 
  geom_point(aes(col=mass), alpha=0.5) + #facet_wrap(~outliers) +
  scale_color_viridis_c("phylogenetic core\nmass")


top_genus_mass = 
  genus_prop %>% 
  tibble::column_to_rownames("taxa") %>% 
  as.matrix %>% 
  apply(1,sum) %>% 
  sort %>% rev %>% head(100)


genus_prop %>%
  filter(taxa %in% c(top_genus_mass %>% names())) %>%
  dplyr::rename(Genus = "taxa") %>%
  #select(1:20) %>%
  #tidyr::separate(taxa, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  #select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  #mutate(Genus = Genus %>% gsub("g__","",.)) %>%
  reshape2::melt(id.vars="Genus") -> top_genus_prop_value_100

metadata %>%
  select("#SampleID","age_cat","age_corrected","sex","country_of_birth","country") %>%
  merge(top_genus_prop_value_100, by.x="#SampleID", by.y="variable") -> top_genus_prop_value_metadata

 
top_genus_prop_value_metadata %>%
  #filter(Genus %in% top_10_genus) %>%
  tidyr::separate(Genus, into=c("Kingdom","Phylum","Class","Order","Family","Genus"), sep=";") %>%
  select(-c("Kingdom","Phylum","Class","Order","Family")) %>%
  mutate(Genus = Genus %>% gsub("g__","",.)) %>% 
  merge(pcoa %>% mutate(pcoa1=as.numeric(pcoa1),pcoa2=as.numeric(pcoa2)) , by.x="#SampleID", by.y="sampleID") %>%
  merge( 
          metadata %>%
  select("#SampleID","age_cat","age_corrected","sex","country_of_birth","country") %>% unique,
  
  by.x="#SampleID", by.y="#SampleID"
  
          )  -> test
          

test %>%
  ungroup() %>%
  group_by(Genus) %>%
  summarise(cor1=cor(value,pcoa1, method="spearman"), cor2=cor(value, pcoa2, method="spearman")) %>%
  arrange(desc(cor2))



```

export outliers
```{r}

test %>%
  filter(Genus %in% c("Acinetobacter","Pseudomonas","Stenotrophomonas")) %>%
  mutate(outliers = ifelse(`#SampleID` %in% microbiota_outliers,"outliers","ok")) %>%
  ggplot(aes(pcoa1, pcoa2)) + geom_point(aes(col=value)) + facet_grid(outliers~Genus) 

test %>%
  filter(Genus %in% c("Acinetobacter","Pseudomonas","Stenotrophomonas")) %>%
  mutate(outliers = ifelse(`#SampleID` %in% microbiota_outliers,"outliers","ok")) %>%
  ggplot(aes(Genus, value+0.0001)) + geom_boxplot(aes(fill=outliers)) + scale_y_log10()

test %>%
  filter(Genus %in% c("Acinetobacter","Pseudomonas","Stenotrophomonas")) %>%
  mutate(outliers = ifelse(`#SampleID` %in% microbiota_outliers,"outliers","ok")) %>%
  filter(value>0.01 | outliers=="outliers") %>%
  #filter(outliers=="outliers") %>%
  pull(`#SampleID`) %>% unique %>% writeLines(con="outliers_samples.txt")



test %>%
  #filter(Genus=="Bifidobacterium") %>%
  #mutate(outliers = ifelse(`#SampleID` %in% microbiota_outliers,"outliers","ok")) %>%
  merge(sample_read_mass, by="#SampleID") %>%
  group_by(Genus) %>%
  summarise(cor_mass=cor(value,mass, method="pearson")) %>%
  arrange((cor_mass))


```



```{r}

outliers = readLines(con="outliers_samples.txt")


test %>%
  filter(Genus %in% c("Acinetobacter","Pseudomonas","Stenotrophomonas")) %>%
  mutate(outliers = ifelse(`#SampleID` %in% outliers,"outliers","ok")) %>%
  ggplot(aes(Genus, value+0.0001)) + geom_boxplot(aes(fill=outliers)) + scale_y_log10()


test %>%
  filter(Genus=="Bifidobacterium") %>%
  mutate(outliers = ifelse(`#SampleID` %in% outliers,"outliers","ok")) %>%
  merge(sample_read_mass, by="#SampleID") %>%
  ggplot(aes(pcoa1, pcoa2)) + 
  geom_point(aes(col=mass), alpha=0.5) + facet_wrap(~outliers) +
  scale_color_viridis_c("phylogenetic core\nmass")

```






